---
title: "RDS Solutions"
author: "Patrick Wilbur"
date: "6/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(ggplot2)) install.packages("ggplot2",repos = "http://cran.us.r-project.org") # install package if not already installed
library(ggplot2) # Load the library
```

# RDS (R for Data Science) Solutions

I know the solutions to these problems already exist on the internet, this is purely for my own benefit

## Chapter 3 Exercises

**Question 1** What’s gone wrong with this code? Why are the points not blue?

```
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

**Answer:** The color should be specified outside of the `aes()` function call, for example

```{r ch3q1, echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

**Question 2** Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

**Answer:** On running `?mpg` you will see the data frame specifications. There are 11 variables in total

*Continues variables*  

* displ  
* cyl  
* cty  
* hwy  

*Categorical variables*  

* manufacturer  
* model  
* year  
* trans  
* drv  
* fl  
* class  

**Question 3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

**Answer:** The following graph shows continuous variables mapped to color and size (shape was not included because it simply errored).

```{r ch3q3, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cyl))
```

The color and size seem to render as a specturm instead of a discrete amount.

