---
title: "RDS Solutions"
author: "Patrick Wilbur"
date: "6/15/2021"
output: md_document
knit: (
  function(inputFile, encoding) { 

    pSubTitle <- 'README'

    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding, 
      params      = list(sub_title = pSubTitle),      
      output_file = pSubTitle) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(ggplot2)) install.packages("ggplot2",repos = "http://cran.us.r-project.org") # install package if not already installed
library(ggplot2) # Load the library
```

# RDS (R for Data Science) Solutions

I know the solutions to these problems already exist on the internet, this is purely for my own benefit

## Chapter 3.2.4 Exercises

**Question 1:** Run ggplot(data = mpg). What do you see?

**Answer:** Not much.

```{r ch3.2.4q1, echo=TRUE}
ggplot(data = mpg)
```

**Question 2:** How many rows are in mpg? How many columns?

**Answer:** 234 rows, 11 columns

```{r ch3.2.4q2, echo=TRUE}
nrow(mpg)
ncol(mpg)
```

**Question 3:** What does the `drv` variable describe? Read the help for ?mpg to find out.

**Answer:**  `the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd`

**Question 4:** Make a scatterplot of hwy vs cyl.

**Answer:** 

```{r ch3.2.4q4, echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = hwy))
```

**Question 5:** What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

**Answer:** Categorical variable vs categorical variable

```{r ch3.2.4q5, echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = class))
```

## Chapter 3.3.1 Exercises

**Question 1** What’s gone wrong with this code? Why are the points not blue?

```
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

**Answer:** The color should be specified outside of the `aes()` function call, for example

```{r ch3.3.1q1, echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

**Question 2** Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

**Answer:** On running `?mpg` you will see the data frame specifications. There are 11 variables in total

*Continues variables*  

* displ  
* cyl  
* cty  
* hwy  

*Categorical variables*  

* manufacturer  
* model  
* year  
* trans  
* drv  
* fl  
* class  

**Question 3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

**Answer:** The following graph shows continuous variables mapped to color and size (shape was not included because it simply errored).

```{r ch3.3.1q3, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cyl))
```

The color and size seem to render on a spectrum instead of a discrete amount.

**Question 4:** What happens if you map the same variable to multiple aesthetics?

**Answer:** The graph below has color and size both mapped to `cty`. Seems to work as expected...

```{r ch3.3.1q4, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cty))
```

**Question 5:** What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

**Answer:** "For shapes that have a border (like 21), you can color the inside and outside separately. Use the stroke aesthetic to modify the width of the border"

**Question 6:** What happens if you map an aesthetic to something other than a variable name, like aes(color = displ < 5)? Note, you’ll also need to specify x and y.

**Answer:** The below graph shows what happens if you set `color = displ < 5`. Since `displ < 5` is a function resulting a boolean, it sets the colors of the graph accordingly 

```{r ch3.3.1q6, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

## Chapter 3.5.1 Exercises

**Question 1:** What happens if you facet on a continuous variable?

**Answer:** Seems to work, but could end up creating a silly amount of graphs that diminish the value of the visualization 

```{r chp3.5.1q1, echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ cty, nrow = 2)
```

**Question 2:** What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

**Answer:** The empty facets in this graph are combinations of drv and cyl that have no data points. For example, the facet "5r" is empty because there are not 5 cyl r drv cars

```{r ch3.5.1q2, echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

**Question 3:** What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

**Answer:** Use the `.` when you prefer to not facet in the rows or columns dimension

```{r ch3.5.1q3, echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

**Question 4:** What are the advantages to using faceting instead of the color aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

**Answer:** Faceting helps show the graph for a specific category. If there are too many categories color would be more appropriate 

```{r ch3.5.1q4, echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

**Question 5:** Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

**Answer:** nrow/ncol specifies the number of rows/columns. scales, shrink, and other arguments exist as well. facet_grid is a grid based of the provided variables, so you can`t specify the number of rows/columns

**Question 6:** When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

**Answer:** This results in more space for the columns

## Chapter 3.6.1 Exercises

**Question 1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

**Answer:** 

* geom_smooth = line
* geom_bar = barplot
* geom_histogram = histogram
* geom_area = area chart

**Question 2:** run the code in R and check your predictions

**Answer:** Ok, my predictions were just about right. I should be an R-runtime in my spare time

```{r ch3.6.1q2, echo=TRUE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

**Question 3:** What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

**Answer:** Removes the legend from the side of the plot, probably to remove clutter

**Question 4:** What does the se argument to geom_smooth() do?

**Answer:** `se` stands for standard error. Setting it true, turns on these gray areas around the line that represent the se

**Question 5:** Will these two graphs look different? Why/why not?

**Answer:** Nope. Same mappings, set globally or locally. 

```{r ch3.6.1q5}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

**Question 6** Recreate the R code necessary to generate the following graphs.

**Answer:**

```{r ch3.6.1q6, echo=TRUE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE, mapping = aes(group = drv))

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE, mapping = aes(group = drv))

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(se = FALSE, mapping = aes(linetype = drv))

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(size = 4, color = "white") +
  geom_point(aes(color = drv))
```